Measuring Hash Trustworthiness via Collision Utility Metrics: Logical Cryptanalysis of MD4

### 2017-12-05 dsn-2018/bare_conf.tex
DSN Figures
DSN progress

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at:
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

@@

Cryptographic hash functions form the core of many protocols. With widespread
use of hash functions as file integrity checks to verify long term storage of
data, as cache invalidation techniques, and as a building block in network
protocols such as Kerberos and TLS, this class of functions necessarily has
strong guarantees about properties of its members. There are three properties
hash functions must have to be considered cryptographically secure:
\begin{itemize}
    \item \textbf{Preimage Resistance}: It should be computationally hard to find
        the inverse of a hash function.
    \item \textbf{Second Preimage Resistance}: Given an input block, it should be
        computationally hard to find a second block which hashes to the
        same value as the first block.
    \item \textbf{Collision Resistance}: It should be computationally hard to find two
        blocks which hash to the same value.
\end{itemize}
Note that while a second preimage is necessarily a collision, it is also a
stronger result with higher use for an attacker.

From an attacker's perspective, finding a preimage or second preimage has been
difficult: the current published bound for a preimage attack in MD5 is a cost
of $2^{123.4}$ by Y. Sasaki and K. Aoki \cite{SasakiPreimage}. In MD4, the
best second preimage attack by H. Yu et al. gives a bound of $2^{-56}$
likelihood, but uses destructive message modification techniques to achieve
collisions for all other messages \cite{cryptoeprint:2007:206}. Further, while
J. Kelsey and B. Schneier proposed a method for finding second preimages
in less than $2^n$, we note that this requires rather long messages
\cite{SchneierSecondPreimage}. However, collision attacks are well within reach
of adversaries, and well past theoretical attacks. The work of
H. Dobbertin \cite{Dobbertin1998}, X. Wang \cite{cryptoeprint:2004:199},
M. Schl{\"a}ffer \cite{Schlaffer2006} and others demonstrate the ease with which
collisions can be found.

From a cryptographic perspective, the existence of a feasible collision attack
breaks the requisite properties, thus deprecating the function's use. In some
instances however, there continues to be widespread use of deprecated hash
functions. One widespread example is the continued use by Git of SHA-1, despite
M. Stevens et al. publishing a full SHA-1 collision in February of 2017
\cite{cryptoeprint:2017:190}. However, this attack had a bound of $2^{63.1}$ --
faster than the theoretical $2^{80}$ but still requiring significant compute
resources to find.

@@

@@

>>>>

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at:
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{wrapfig}

@@

Cryptographic hash functions form the core of many protocols which
trust that cryptographic hash functions preserve certain basic
properties. Cryptograpic hash functions see widespread
use for file integrity checks to verify long term storage of
data, as cache invalidation techniques, and as a building block in network
protocols such as Kerberos and TLS.  The classes of functions suitable
for these use cases necessarily have
strong guarantees about properties of its members. 
An ideal cryptographic hash has five main features: determinism - the
same message always results in the same hash; performability -
computing a hash for a message should not be algorithmically
difficult; one-way - it should be infeasible to generate a message
from its hash except by trying all possible messages; sensitive to
initial conditions - small changes to a message should change the hash
value to a degree that the new hash value appears uncorrelated with
the old hash value; unique digests - it should be infeasible to find
two different messages with the same hash value.

The most important aspect of these can be formally defined as three
properties which hash functions must have to be considered cryptographically secure:
\begin{itemize}
    \item \textbf{Preimage Resistance}: It should be computationally hard to find
        the inverse of a hash function.
    \item \textbf{Second Preimage Resistance}: Given an input block, it should be
        computationally hard to find a second block which hashes to the
        same value as the first block.
    \item \textbf{Collision Resistance}: It should be computationally hard to find two
        blocks which hash to the same value.
\end{itemize}
Note that while a second preimage is necessarily a collision, it is also a
stronger result with higher use for an attacker.

From an attacker's perspective, finding a preimage or second preimage has been
difficult: the current published bound for a preimage attack in MD5 is a cost
of $2^{123.4}$ by Y. Sasaki and K. Aoki \cite{SasakiPreimage}. In MD4, the
best second preimage attack by H. Yu et al. gives a bound of $\frac{1}{2^{56}}$
likelihood, but uses destructive message modification techniques to achieve
collisions for all other messages \cite{cryptoeprint:2007:206}. Further, while
J. Kelsey and B. Schneier proposed a method for finding second preimages
in less than $2^n$, we note that this requires rather long messages
\cite{SchneierSecondPreimage} which limits their utility to
attackers. Collision attacks, however, are well within reach
of adversaries, and have progressed past theoretical attacks into feasibility. The work of
H. Dobbertin \cite{Dobbertin1998}, X. Wang \cite{cryptoeprint:2004:199},
M. Schl{\"a}ffer \cite{Schlaffer2006} and others demonstrate the ease with which
collisions can be found.

\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{Figs/graph-intro-crop.pdf}
\caption{A visualization of the heuristic unit-distance neighborhood
  of valid collisions from the Dobbertin and Kasselman collisions.}
\label{Fig:Graph-Intro}
\end{center}
\end{figure}

From a cryptographic perspective, the existence of a feasible collision attack
breaks the requisite properties, thus deprecating the function's use
for trustworthy applications. In many practical instances, however,
communities continue to use deprecated hash functions in a widespread fashion.
One such example of widespread use post-deprecation is the continued use by Git of SHA-1, despite
M. Stevens et al. publishing a full SHA-1 collision in February of 2017
\cite{cryptoeprint:2017:190}. This attack had a bound of $2^{63.1}$
\emdash faster than the theoretical $2^{80}$ but still requiring significant compute
resources to find.

@@

\begin{figure*}
\begin{center}
  \begin{subfigure}{\textwidth}
    \includegraphics[width=\textwidth]{Figs/differential-1289.png}
    \caption{Collision 1289}
    \label{Fig:1289}
  \end{subfigure}\\
  \begin{subfigure}{\textwidth}
    \includegraphics[width=\textwidth]{Figs/differential-1250.png}
    \caption{Collision 1250}
    \label{Fig:1250}
  \end{subfigure}\\
  \begin{subfigure}{\textwidth}
    \includegraphics[width=\textwidth]{Figs/transitions.png}
    \caption{Differential Path Transitions}
    \label{Fig:Transitions}
  \end{subfigure}\\
\end{center}
\end{figure*}

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{Figs/graph-clique-crop.pdf}
\caption{}
\label{Fig:Graph-Clique}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{Figs/graph-neighborhood-crop.pdf}
\caption{}
\label{Fig:Graph-Neighborhood}
\end{center}
\end{figure}
